---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r loading data pred year, warning=FALSE, message=TRUE, results='asis'}
data_beans = read.csv("data/DataBean_MET_GYv2.csv",h=T, stringsAsFactors = T)

# Data adjustment
# All the effect columns must be as a factor to run in ASReml-r.
cols <- c("rep", "name", "loc","year", "mkt", "year_loc")
data_beans[cols] <- lapply(data_beans[cols], factor)
data_beans <- data.table(data_beans)

```

### Box plot dist.

```{r Box plot_raw data complete, eval= FALSE, warning=FALSE, message=FALSE, fig.heiht=figheight, fig.width=figwidth, dpi = 600, out.width="100%"}

plotDM_stats_mkt_loc<- grouped_ggbetweenstats(data=data_beans, x= loc, y=gy_kg_ha, type = "parametric",  bf.message = F, results.subtitle = F, 
                              ylab= "GY", xlab = "Locations",
                              plot.type = "boxviolin", grouping.var = mkt ) 
#print(plotDM_stats_mkt_loc)

plotDM_stats_mkt_year<- grouped_ggbetweenstats(data=data_beans, x= year, y=gy_kg_ha, type = "parametric",  bf.message = F, results.subtitle = F, 
                              ylab= "GY", xlab = "Years",
                              plot.type = "boxviolin", grouping.var = mkt ) 
#print(plotDM_stats_mkt_year)

print(arrange_ggplot(plotDM_stats_mkt_loc,plotDM_stats_mkt_year))

```


```{r Box plot_raw data, eval=TRUE, fig.heiht=figheight, fig.width=figwidth, message=FALSE, warning=FALSE, dpi=600, out.width="100%"}

a<- ggplot(data=data_beans, aes(x=reorder(loc, -gy_kg_ha), y=gy_kg_ha, fill=loc)) +
  ggdist::stat_halfeye(
    adjust = 0.5,
    justification = -0.1,
    .width = 0,
    point_colour = NA
  ) +
  
  geom_boxplot( 
    width = .12,
   # outlier.color = NA,
    alpha = 0.5)+
  tidyquant::theme_tq()+
  facet_wrap("year")+
  facet_grid("mkt") +
  tidyquant::scale_fill_tq() +
  scale_x_discrete(expand = c(0.03,0)) +
  scale_y_continuous(limits = c(1000,6000), breaks = seq(1000, 6000, by = 1000)) +
  theme(axis.text.x=element_text(angle = 90),
        strip.text=element_blank(),
        legend.position = "none",
        panel.grid = element_blank())+
  labs(title="GY (kg/ha)", 
       subtitle="Data distribution by location", 
       caption=NULL, x=NULL, y=NULL)

b<- ggplot(data=data_beans, aes(x=reorder(year, -gy_kg_ha), y=gy_kg_ha, fill=year)) +
  ggdist::stat_halfeye(
    adjust = 0.5,
    justification = -0.1,
    .width = 0,
    point_colour = NA
  ) +
  
  geom_boxplot( 
    width = .12,
    #outlier.color = NA,
    alpha = 0.5)+
  tidyquant::theme_tq()+
  facet_wrap("loc")+
  facet_grid("mkt") +
  tidyquant::scale_fill_tq() +
  scale_x_discrete(expand = c(0.03,0)) +
  scale_y_continuous(limits = c(1000,6000), breaks = seq(1000, 6000, by = 1000)) +
  theme(axis.text.x=element_text(angle = 90),
        strip.text=element_text(face="bold"),
        legend.position = "none",
        axis.text.y = element_blank(),
         panel.grid = element_blank())+
  labs(title="GY (kg/ha)", 
       subtitle="Data distribution by year", 
       caption=NULL, x=NULL, y=NULL)

print(arrange_ggplot(a,b))

```

#### BB
```{r}

data_beans_plotBB<- droplevels(subset(data_beans, mkt=="BB"))

stats_results_BB<- ggbetweenstats(data=data_beans_plotBB, x= loc, y=gy_kg_ha) %>% extract_stats()   
#print(stats_results_BB$pairwise_comparisons_data$p.value)

stats_results_BB2<- stats_results_BB$pairwise_comparisons_data %>% 
dplyr::mutate(groups = purrr::pmap(.l = list(group1, group2), .f = c)) %>%
  dplyr::arrange(group1)%>% 
  filter(p.value <= 0.05)

bb1<- ggplot(data=data_beans_plotBB, aes(x=reorder(loc, -gy_kg_ha), y=gy_kg_ha, fill=loc)) +
  ggdist::stat_halfeye(
    adjust = 0.5,
    justification = -0.1,
    .width = 0,
    point_colour = NA
  ) +
  
  geom_boxplot( 
    width = .12,
   # outlier.color = NA,
    alpha = 0.5)+
  tidyquant::theme_tq()+
  #facet_grid("loc")+
  tidyquant::scale_fill_tq() +
  scale_x_discrete(expand = c(0.03,0)) +
  scale_y_continuous(limits = c(1000,10000), breaks = seq(1000, 10000, by = 1000)) +
  theme(axis.text.x=element_blank(),
        strip.text=element_blank(),
        legend.position = "none",
        panel.grid = element_blank())+
  labs(title="GY (kg/ha)", 
       subtitle="Data distribution by location", 
       caption=NULL, x=NULL, y="BB")

bb1<-bb1 +
  ggsignif::geom_signif(
    comparisons      = stats_results_BB2$groups,
    map_signif_level = TRUE,
    tip_length       = 0.01,
    textsize         = 2.7,
    y_position       = c(5900, 6700, 7500,8300, 9100),
    annotations      = as.character(stats_results_BB2$expression),
    test             = NULL,
    na.rm            = TRUE,
    parse            = TRUE
  )
```

```{r}
data_beans_plotBB<- droplevels(subset(data_beans, mkt=="BB"))

stats_results_BB<- ggbetweenstats(data=data_beans_plotBB, x= year, y=gy_kg_ha) %>% extract_stats()   
#print(stats_results_BB$pairwise_comparisons_data$p.value)

stats_results_BB3<- stats_results_BB$pairwise_comparisons_data %>% 
dplyr::mutate(groups = purrr::pmap(.l = list(group1, group2), .f = c)) %>%
  dplyr::arrange(group1) %>% 
  filter(p.value <= 0.05)

bb2<- ggplot(data=data_beans_plotBB, aes(x=reorder(year, -gy_kg_ha), y=gy_kg_ha, fill=year)) +
  ggdist::stat_halfeye(
    adjust = 0.5,
    justification = -0.1,
    .width = 0,
    point_colour = NA
  ) +
  
  geom_boxplot( 
    width = .12,
   # outlier.color = NA,
    alpha = 0.5)+
  tidyquant::theme_tq()+
  #facet_grid("loc")+
  tidyquant::scale_fill_tq() +
  scale_x_discrete(expand = c(0.03,0)) +
  scale_y_continuous(limits = c(1000,10000), breaks = seq(1000, 10000, by = 1000)) +
  theme(axis.text.x=element_blank(),
        strip.text=element_blank(),
        legend.position = "none",
        panel.grid = element_blank())+
  labs(title="GY (kg/ha)", 
       subtitle="Data distribution by year", 
       caption=NULL, x=NULL, y=NULL)


bb2<-bb2 +
  ggsignif::geom_signif(
    comparisons      = stats_results_BB3$groups,
    map_signif_level = TRUE,
    tip_length       = 0.01,
    textsize         = 2.7,
    y_position       = c(5900, 6300, 6700,7100, 7400, 7800,8200, 8600, 9000, 9400),
    annotations      = as.character(stats_results_BB3$expression),
    test             = NULL,
    na.rm            = TRUE,
    parse            = TRUE
  )
```


#### NB
```{r}

data_beans_plotBB<- droplevels(subset(data_beans, mkt=="NB"))

stats_results_BB<- ggbetweenstats(data=data_beans_plotBB, x= loc, y=gy_kg_ha) %>% extract_stats()   
#print(stats_results_BB$pairwise_comparisons_data$p.value)

stats_results_BB2<- stats_results_BB$pairwise_comparisons_data %>% 
dplyr::mutate(groups = purrr::pmap(.l = list(group1, group2), .f = c)) %>%
  dplyr::arrange(group1)%>% 
  filter(p.value <= 0.05)

nb1<- ggplot(data=data_beans_plotBB, aes(x=reorder(loc, -gy_kg_ha), y=gy_kg_ha, fill=loc)) +
  ggdist::stat_halfeye(
    adjust = 0.5,
    justification = -0.1,
    .width = 0,
    point_colour = NA
  ) +
  
  geom_boxplot( 
    width = .12,
   # outlier.color = NA,
    alpha = 0.5)+
  tidyquant::theme_tq()+
  #facet_grid("loc")+
  tidyquant::scale_fill_tq() +
  scale_x_discrete(expand = c(0.03,0)) +
  scale_y_continuous(limits = c(1000,10000), breaks = seq(1000, 10000, by = 1000)) +
  theme(axis.text.x=element_blank(),
        strip.text=element_blank(),
        legend.position = "none",
        panel.grid = element_blank())+
  labs( 
       caption=NULL, x=NULL, y="NB")

nb1<-nb1 +
  ggsignif::geom_signif(
    comparisons      = stats_results_BB2$groups,
    map_signif_level = TRUE,
    tip_length       = 0.01,
    textsize         = 2.7,
    y_position       = c(5900, 6700, 7500,8300, 9100),
    annotations      = as.character(stats_results_BB2$expression),
    test             = NULL,
    na.rm            = TRUE,
    parse            = TRUE
  )
```

```{r}
data_beans_plotBB<- droplevels(subset(data_beans, mkt=="NB"))

stats_results_BB<- ggbetweenstats(data=data_beans_plotBB, x= year, y=gy_kg_ha) %>% extract_stats()   
#print(stats_results_BB$pairwise_comparisons_data$p.value)

stats_results_BB3<- stats_results_BB$pairwise_comparisons_data %>% 
dplyr::mutate(groups = purrr::pmap(.l = list(group1, group2), .f = c)) %>%
  dplyr::arrange(group1) %>% 
  filter(p.value <= 0.05)

nb2<- ggplot(data=data_beans_plotBB, aes(x=reorder(year, -gy_kg_ha), y=gy_kg_ha, fill=year)) +
  ggdist::stat_halfeye(
    adjust = 0.5,
    justification = -0.1,
    .width = 0,
    point_colour = NA
  ) +
  
  geom_boxplot( 
    width = .12,
   # outlier.color = NA,
    alpha = 0.5)+
  tidyquant::theme_tq()+
  #facet_grid("loc")+
  tidyquant::scale_fill_tq() +
  scale_x_discrete(expand = c(0.03,0)) +
  scale_y_continuous(limits = c(1000,10000), breaks = seq(1000, 10000, by = 1000)) +
  theme(axis.text.x=element_blank(),
        strip.text=element_blank(),
        legend.position = "none",
        panel.grid = element_blank())+
  labs( 
       caption=NULL, x=NULL, y=NULL)


nb2<-nb2 +
  ggsignif::geom_signif(
    comparisons      = stats_results_BB3$groups,
    map_signif_level = TRUE,
    tip_length       = 0.01,
    textsize         = 2.7,
    y_position       = c(5900, 6300, 6700,7100, 7400, 7800,8200, 8600, 9000, 9400),
    annotations      = as.character(stats_results_BB3$expression),
    test             = NULL,
    na.rm            = TRUE,
    parse            = TRUE
  )
```

#### SR
```{r}

data_beans_plotBB<- droplevels(subset(data_beans, mkt=="SR"))

stats_results_BB<- ggbetweenstats(data=data_beans_plotBB, x= loc, y=gy_kg_ha) %>% extract_stats()   
#print(stats_results_BB$pairwise_comparisons_data$p.value)

stats_results_BB2<- stats_results_BB$pairwise_comparisons_data %>% 
dplyr::mutate(groups = purrr::pmap(.l = list(group1, group2), .f = c)) %>%
  dplyr::arrange(group1)%>% 
  filter(p.value <= 0.05)

sr1<- ggplot(data=data_beans_plotBB, aes(x=reorder(loc, -gy_kg_ha), y=gy_kg_ha, fill=loc)) +
  ggdist::stat_halfeye(
    adjust = 0.5,
    justification = -0.1,
    .width = 0,
    point_colour = NA
  ) +
  
  geom_boxplot( 
    width = .12,
   # outlier.color = NA,
    alpha = 0.5)+
  tidyquant::theme_tq()+
  #facet_grid("loc")+
  tidyquant::scale_fill_tq() +
  scale_x_discrete(expand = c(0.03,0)) +
  scale_y_continuous(limits = c(1000,10000), breaks = seq(1000, 10000, by = 1000)) +
  theme(
        strip.text=element_blank(),
        legend.position = "none",
        panel.grid = element_blank())+
  labs(
       caption=NULL, x="Locations", y="SR")

sr1<-sr1 +
  ggsignif::geom_signif(
    comparisons      = stats_results_BB2$groups,
    map_signif_level = TRUE,
    tip_length       = 0.01,
    textsize         = 2.7,
    y_position       = c(5900, 6700, 7500,8300, 9100),
    annotations      = as.character(stats_results_BB2$expression),
    test             = NULL,
    na.rm            = TRUE,
    parse            = TRUE
  )
```

```{r}
data_beans_plotBB<- droplevels(subset(data_beans, mkt=="SR"))

stats_results_BB<- ggbetweenstats(data=data_beans_plotBB, x= year, y=gy_kg_ha) %>% extract_stats()   
#print(stats_results_BB$pairwise_comparisons_data$p.value)

stats_results_BB3<- stats_results_BB$pairwise_comparisons_data %>% 
dplyr::mutate(groups = purrr::pmap(.l = list(group1, group2), .f = c)) %>%
  dplyr::arrange(group1) %>% 
  filter(p.value <= 0.05)

sr2<- ggplot(data=data_beans_plotBB, aes(x=reorder(year, -gy_kg_ha), y=gy_kg_ha, fill=year)) +
  ggdist::stat_halfeye(
    adjust = 0.5,
    justification = -0.1,
    .width = 0,
    point_colour = NA
  ) +
  
  geom_boxplot( 
    width = .12,
   # outlier.color = NA,
    alpha = 0.5)+
  tidyquant::theme_tq()+
  #facet_grid("loc")+
  tidyquant::scale_fill_tq() +
  scale_x_discrete(expand = c(0.03,0)) +
  scale_y_continuous(limits = c(1000,10000), breaks = seq(1000, 10000, by = 1000)) +
  theme(
        strip.text=element_blank(),
        legend.position = "none",
        panel.grid = element_blank())+
  labs(
       caption=NULL, y=NULL, x="Years")


sr2<- sr2 +
  ggsignif::geom_signif(
    comparisons      = stats_results_BB3$groups,
    map_signif_level = TRUE,
    tip_length       = 0.01,
    textsize         = 2.7,
    y_position       = c(5900, 6500, 7200,8000, 8700, 9500),
    annotations      = as.character(stats_results_BB3$expression),
    test             = NULL,
    na.rm            = TRUE,
    parse            = TRUE
  )
```


Box plot distribution for each market class in this study across years and locations.
```{r plot_Box complete final, warning=FALSE, message=FALSE, fig.heiht=figheight, fig.width=figwidth, dpi = 600, out.width="100%"}

print(arrange_ggplot(bb1,bb2,
                     nb1,nb2,
                     sr1,sr2,
                     nrow = 3,
                     ncol = 2))

```


### GEI Comparisons (loc:year) - BB 

Box plots for between-subjects comparisons by locations using the R package `ggstatsplot`.

- Black beans (BB)
**Pairwise Games-Howell test used. Comparisons showing only significant**

```{r plot_ggstatsplot_desc_BB, warning=FALSE, message=FALSE, fig.heiht=figheight, fig.width=figwidth, dpi = 600, out.width="100%"}
data_beans_plotBB<- droplevels(subset(data_beans, mkt=="BB"))

plotDM_stats_mkt_loc_BB<- grouped_ggbetweenstats(data=data_beans_plotBB, x= loc, y=gy_kg_ha, type = "parametric",  bf.message = F, results.subtitle = F, 
                              ylab= "GY", xlab = "Locations",
                              plot.type = "boxviolin", grouping.var = year ) 
print(plotDM_stats_mkt_loc_BB)

```



### GEI Comparisons (loc:year) - NB 

Box plots for between-subjects comparisons by locations using the R package `ggstatsplot`.
- Navy beans (NB)
**Pairwise Games-Howell test used. Comparisons showing only significant**

```{r plot_ggstatsplot_desc_NB, warning=FALSE, message=FALSE, fig.heiht=figheight, fig.width=figwidth, dpi = 600, out.width="100%"}

data_beans_plotNB<- droplevels(subset(data_beans, mkt=="NB"))

plotDM_stats_mkt_loc_NB<- grouped_ggbetweenstats(data=data_beans_plotNB, x= loc, y=gy_kg_ha, type = "parametric",  bf.message = F, results.subtitle = F, 
                              ylab= "GY", xlab = "Locations",
                              plot.type = "boxviolin", grouping.var = year ) 
print(plotDM_stats_mkt_loc_NB)
```

### GEI Comparisons (loc:year) - SR 

Box plots for between-subjects comparisons by locations using the R package `ggstatsplot`.
- Black beans (SR)
**Pairwise Games-Howell test used. Comparisons showing only significant**

```{r plot_ggstatsplot_desc_SR, warning=FALSE, message=FALSE, fig.heiht=figheight, fig.width=figwidth, dpi = 600, out.width="100%"}

data_beans_plotSR<- droplevels(subset(data_beans, mkt=="SR"))

plotDM_stats_mkt_loc_SR<- grouped_ggbetweenstats(data=data_beans_plotSR, x= loc, y=gy_kg_ha, type = "parametric",  bf.message = F, results.subtitle = F, 
                              ylab= "GY", xlab = "Locations",
                              plot.type = "boxviolin", grouping.var = year ) 
print(plotDM_stats_mkt_loc_SR)
```


### Scatter plot

For further insight into the correlation structure between trials a scatter plot matrix can be made using the package `statgenGxE`

```{r TDscatter2, warning=FALSE, out.width="100%", message=FALSE, fig.heiht=figheight, fig.width=figwidth, fig.align="center", dpi = 600}
dropsTD <- statgenSTA::createTD(data = data_beans, genotype = "name", trial = "loc")
options("statgen.genoColors" = c("black", "blue", "red"))
plot(dropsTD, plotType = "scatter", traits = "gy_kg_ha", colorGenoBy = "mkt", 
     colorTrialBy = "trial", title = "Scatterplots of trials for grain yield (Kg/ha)" )

```


### Outliers

Check for common errors in multi-environment trial by market class data using the R package `metan`
However, the data set was cleaned before at a previous analysis not shown in this vignette using the criteria for GY:

- yield <= 0.5 | yield >= 7.5 (considering yield as Lb per plot)
- moisture >= 30 | moisture <= 5 (considering moisture in %)

At this point yield data will not be removed anymore, but we can check it out for further investigation, if it was needed. 

```{r data_outliers1, warning=FALSE, message=TRUE, results='asis'}

out_beans<- find_outliers(data_beans, by=mkt, var = gy_kg_ha, plots = F)

#out_beans
```

- No outliers removed after the data inspection based on the Breeder expertise.

### Map of locations 

This figure represent all counties locations evaluated in this study at Michigan, midwest in the USA.

```{r locations map, warning=FALSE, out.width="100%", message=FALSE, fig.heiht=figheight, fig.width=figwidth, fig.align="center", dpi = 600, fig.pos="H"}

source("utils/map.R")

print(MI_trials_beans_map)

```


### Missing values

```{r missing geno gy, warning=FALSE, message=FALSE, out.width="100%", fig.pos="H", fig.align="center", dpi = 600, fig.cap="Missing genotypes by market classes"}

data_beans.miss.f<- data_beans %>% 
  unite(name_year_loc, c(name, year_loc), remove = F) %>% 
  group_by(name_year_loc) %>% 
  dplyr::summarise(Mean = mean(gy_kg_ha, na.rm = TRUE)) %>% 
  filter(!is.na(Mean)) 

data_beans.miss<- data_beans %>% 
  unite(name_year_loc, c(name, year_loc), remove = F)

data_beans.miss<- data_beans.miss%>% 
  filter((name_year_loc %in% data_beans.miss.f$name_year_loc))  

gg_miss_fct(data_beans.miss,fct = mkt)+ggtitle("Missing values (NA)") + 
   theme(axis.text.x=element_text(face="bold", size = 12), 
         axis.text.y = element_text(face = "bold",size = 12) ,
         title = element_text(face = "bold",size = 14))
```



### Coincidence Genotypes

> Genotypes coincidentes across years combinations

```{r coincidence geno fig1, warning=FALSE,  fig.pos="H", message=FALSE, out.width="100%", fig.align="center", dpi = 600, fig.cap="Coincidence genotypes per year and combinations of years across locations for each market class."}

source("utils/Coinc.R")

arrange_ggplot(plotB, plotN, plotSR,ncol = 1)

```













