---
output: html_document
editor_options: 
  chunk_output_type: console
---

### MET analysis - ASReml

Running MET using `ASReml` - only to comparison of variance components with `metan` outputs

```{r mod met asreml all_kt, warning=FALSE, message=TRUE, results='asis'}

mod.met.asreml <- asreml(fixed = yield ~ loc + loc:rep,
                random = ~ name + name:loc,
                data = blues_stage.I,
                predict = predict.asreml(classify = "name"),
                trace = F,
                maxit = 500)
summFix.met.asreml<- data.frame(wald(mod.met.asreml))
summFix.met.asreml

summ.met.asreml<- data.frame(summary.asreml(mod.met.asreml)$varcomp)
summ.met.asreml

#print(summary.asreml(mod.met.asreml)$bic)
#mod.met.asreml<- data.table((mod.met.asreml$predictions$pvals[1:3]))
mod.met.asreml2 <- data.table(matrix(unlist(mod.met.asreml$predictions$pvals[1:3]),
 ncol=3))
names(mod.met.asreml2) <- c("name", "yield_BLUPS_MET", "SE")
mod.met.asreml2$yield_BLUPS_MET<-as.numeric(mod.met.asreml2$yield_BLUPS_MET)
mod.met.asreml2$SE<-as.numeric(mod.met.asreml2$SE)

```


### MET analysis - lme4

Running MET using `metan` R package [@olivotoMeanPerformanceStability2019](Olivoto et al., 2019a).

```{r mod metan, warning=FALSE, message=TRUE, results='asis'}
mixed_mod<- 
   gamem_met(blues_stage.I,
             env = loc,
             gen = name,
             rep = rep,
             resp = yield,
             random = "gen", #Default
             verbose = TRUE) #Default

```

### Printing the model outputs {.tabset}

#### Likelihood Ratio Tests

-   The output `LRT` contains the Likelihood Ratio Tests for genotype and genotype-vs-environment random effects. 

```{r  warning=FALSE, message=TRUE, results='asis'}
data_mod_bb_test <- get_model_data(mixed_mod, "lrt")
data_mod_bb_test

#customize the display of numbers and other data in a tibble
# old <- options(pillar.sigfig = 6)
# 
# blues_stage.I %>% 
#   group_by(loc) %>% 
#   dplyr::summarise(Mean = mean(yield, na.rm = TRUE))

```

#### Detailed parameters

```{r  warning=FALSE, message=TRUE, results='asis'}
data_mod_bb_det <- get_model_data(mixed_mod, "details")
data_mod_bb_det

```



#### Random effects 

-   The output `LRT` contains the Likelihood Ratio Tests for genotype and genotype-vs-environment random effects. 

```{r  warning=FALSE, message=TRUE, results='asis'}
#customize the display of numbers and other data in a tibble
old <- options(pillar.sigfig = 8)

data_mod_var <- get_model_data(mixed_mod, "vcomp")
data_mod_var
```


#### Variance components and genetic parameters 


```{r warning=FALSE, message=TRUE, results='asis' }

old <- options(pillar.sigfig = 4)
data_mod_comp <- get_model_data(mixed_mod)
data_mod_comp
```


### MET - GGE biplot {.tabset}

Genotype plus Genotype-vs-Environment interaction (GGE). Mega-environment identification in multi-environment trials (MET) according to W. Yan et al. 2007.

#### GGE ENV biplot


GGE biplot done using: 

- **sd**: each value is divided by the standard deviation of its corresponding environment.
- **environment**: environment-centered (G+GE)
- **environment**: singular value is entirely partitioned into the environment eigenvectors, also called column metric preserving


```{r gge fig1, warning=FALSE, message=FALSE, fig.height=figheight, fig.width=figwidth, fig.align="center", dpi = 600}
gge_model <- gge(blues_stage.I, loc, name, yield, 
                    centering = "environment", #2
                    scaling = "sd", #2
                    svp = "environment")#2

a <- plot(gge_model, type=4,
          size.text.env = 4.5,
          plot_theme = theme_metan(grid =  "both",color.background = transparent_color()),
         axis_expand = 1.5,
         col.alpha.circle = 0.8,
         shape.gen = NA,
         col.gen = NA,
         size.text.lab = NA,
         size.text.gen = NA,
         leg.lab=c('Env'),
        #title = FALSE
         )


gge_model <- gge(blues_stage.I, loc, name, yield,
                    centering = "environment", #1
                    scaling = "sd", #2Y
                    svp = "environment")#2)

b <- plot(gge_model, type = 6,
          size.text.env = 5,
          plot_theme = theme_metan(grid =  "both",color.background = transparent_color()),
         axis_expand = 1.5,
        # col.alpha.circle = 100,
          col.alpha.circle = 0.8,
         size.text.lab = 13,
       #title = FALSE
        )

 arrange_ggplot(a, b,
                guides = "collect",
  tag_levels = "a",
  tag_prefix = "(",
  tag_suffix = ")")
 
 
```


#### Biplot type 3: Which-won-where

GGE biplot done using: 

- **sd**: each value is divided by the standard deviation of its corresponding environment.
- **environment**: environment-centered (G+GE)
- **genotype**: singular value is entirely partitioned into the environment eigenvectors, also called column metric preserving


```{r gge fig2, warning=FALSE, message=FALSE, fig.height=figheight, fig.width=figwidth, fig.align="center", dpi = 600}
gge_model <- gge(blues_stage.I, loc, name, yield,
                     centering = "environment", #2
                    scaling = "sd", #1
                    svp = "genotype")#2)

e <- plot(gge_model, type = 3,
          size.text.env = 5,
          plot_theme = theme_metan(grid =  "both",color.background = transparent_color()),
         axis_expand = 1.2,
         size.line = 0.7,
         size.text.gen = 4,
         size.text.win = 4.5
         #title = F
        )

print(e)

```


### Mean performance and stability analysis {.tabset}

WAASP index and BLUPs to estimate stability analysis.

```{r mod metan stab, warning=FALSE, message=FALSE, fig.height=figheight, fig.width=figwidth, fig.align="center", dpi = 600}
#blues_stage.I<- na.omit(blues_stage.I)

waasb_model_allMKT <- 
  waasb(blues_stage.I,
        env = loc,
        gen = name,
        rep = rep,
        resp = yield,
        random = "gen", #Default
        verbose = TRUE,
        wresp = 60) #weight for response variable 60 and 40 for yielding and stability, respectively)

waasb_model<- waasb_model_allMKT$yield$model

#waasb_ind <- gmd(waasb_model, "WAASB")
#print_tbl(waasb_ind)

#desc <- c("Selected cultivar providing greater performance and stability for GY")

waasp_plot <- plot_scores(waasb_model_allMKT, type = 3,
          title = FALSE,
          size.tex.gen = 4,
          size.tex.env = 4,
          size.tex.lab = 13,
        # highlight = c("B55", "B1" , "B29", "B20" ,"B28"),
         plot_theme = theme_metan(grid =  "both",color.background = transparent_color())
        ) +
  
  geom_mark_rect(aes(filter =  Code  %in% c("N64", "N40", "B68", "N7"),
                     ),
               label.fontsize = 10,
               show.legend = F,
               con.cap = 0,
               con.colour = "red",
               color = "red",
               expand = 0.01,
               label.buffer = unit(10, "cm"))+
#theme_gray()+
theme(legend.position = c(0.1, 0.9),
      legend.background = element_blank(),
      legend.title = element_blank(),
      aspect.ratio = 1) +
  labs(x = "GY") 

print(waasp_plot)

```

```{r selected waasb22, warning=FALSE, message=TRUE, results='asis'}

waasb_model_meanWaasb<-mean(waasb_model$WAASB)
waasb_model_meanY<-mean(waasb_model$Y)

selected <- waasb_model %>%
  dplyr::filter(Y >= waasb_model_meanY & WAASB <= waasb_model_meanWaasb) 

selected_table <- selected

if (knitr::is_html_output()) {

  print_table(selected_table)
  
}else{
  
 selected_table[,1:8] 
}


#selected$Code

```


#### Selection differentials

```{r mod metan stab2, warning=FALSE, message=FALSE, fig.height=figheight, fig.width=figwidth, fig.align="center", dpi = 600, fig.cap="Mean performance for grain yield (GY) of all dry beans genotypes present in this study colored by market classes. The vertical dashed and solid lines shows, respectivelly, the mean of the selected genotype and the overall mean for both mean performance and WAASB index"}

#Create a data frame with BLUPS - selected and non-selected
blups_sel <-
  gmd(waasb_model_allMKT, "blupge") %>%
  add_cols(SELECTED = ifelse(GEN %in% selected$Code, "yes", "no")) %>% 
    dplyr::rename(BLUPs_sel = yield) %>% 
  droplevels()

blups_sel_mean<-
  gmd(waasb_model_allMKT, "blupge") %>%
  add_cols(SELECTED = ifelse(GEN %in% selected$Code, "yes", "no")) %>% 
  filter(SELECTED == "yes") %>% 
  dplyr::summarise(mean_GY = mean(yield,na.rm = TRUE), n = n()) 

# Create a data frame with the waasb index - selected and non-selected
waasb_sel <-
  gmd(waasb_model_allMKT, "WAASB") %>%
  add_cols(SELECTED = ifelse(GEN %in% selected$Code, "yes", "no")) %>% 
  dplyr::rename(WAASB_sel = yield) %>% 
  droplevels()
#str(waasb_sel)

waasb_sel_mean<-
  gmd(waasb_model_allMKT, "WAASB") %>%
  add_cols(SELECTED = ifelse(GEN %in% selected$Code, "yes", "no")) %>% 
  filter(SELECTED == "yes") %>% 
  dplyr::summarise(mean_GY = mean(yield,na.rm = TRUE), n = n()) 

p1<- plot_selected(blups_sel, GEN, BLUPs_sel, mean_sel = blups_sel_mean$mean_GY) +
  labs(y = "GY") 


blups_sel2<- blups_sel %>% 
  dplyr::filter(SELECTED == "yes") 

blups_sel2$GEN <- factor(blups_sel2$GEN, levels = unique(blups_sel2$GEN))
blups_sel2$color <- ifelse(str_starts(blups_sel2$GEN, "B"), "black",
                           ifelse(str_starts(blups_sel2$GEN, "R"), "red", "blue"))

blups_sel2 <- blups_sel2 %>%
  dplyr::arrange(color, desc(BLUPs_sel)) %>%
  dplyr::group_by(color) %>% 
  dplyr::mutate(order = row_number())


p11 <-
 ggplot(blups_sel2, aes(x = GEN, y = BLUPs_sel, fill = color, alpha = 0.6)) +  # Use color column for fill aesthetic
  stat_summary(fun = mean,
               geom = "bar",
               na.rm = TRUE,
               color = "black",
               size = 0.1,
               width = 1) +
  stat_summary(fun.data = mean_se,
               geom = "errorbar",
               na.rm = TRUE,
               color = "black",
               size = 0.1,
               width = .5) +
  theme_bw() +
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  theme(panel.grid = element_blank(),
        axis.text = element_text(size = 8, colour = "black"),
        axis.text.y = element_text(size = 8, colour = "black"),
        legend.position = "",
        axis.text.x = element_text(angle = 90, color = blups_sel2$color)) +
  geom_hline(yintercept = mean(blups_sel$BLUPs_sel, na.rm = TRUE), linetype = 1) +
  geom_hline(yintercept = blups_sel_mean$mean_GY, linetype = 2) +
  labs(x = "Genotype",
       y = "GY") +
  coord_flip() +
  scale_fill_identity() +
  scale_color_identity() 
  
  print(p11)

```




```{r mod metan stab3, warning=FALSE, message=FALSE, fig.height=figheight, fig.width=figwidth, fig.align="center", dpi = 600, fig.cap="Mean performance (a) and stability (b) for grain yield (GY) for all dry beans genotypes present in this study. The vertical dashed and solid lines shows, respectivelly, the mean of the selected genotype and the overall mean for both mean performance and WAASB index"}

p3<- plot_selected(waasb_sel, GEN, WAASB_sel, mean_sel = waasb_sel_mean$mean_GY) +
  labs(y = "WAASB index")

arrange_ggplot(p1, p3,
  guides = "collect",
  tag_levels = "a",
  tag_prefix = "(",
  tag_suffix = ")")


```



Percentage (SD_gain in %) gain from the selected genotypes compared to the general mean.

```{r mod metan stab3_gain, warning=FALSE, message=TRUE, results='asis'}
blups_sel2 <-
  gmd(waasb_model_allMKT, "blupg") %>%
  add_cols(SELECTED = ifelse(GEN %in% selected$Code, "yes", "no")) %>% 
    dplyr::rename(BLUPs_sel = yield) %>% 
  droplevels()

blups_sel_mean2<-
  gmd(waasb_model_allMKT, "blupg") %>%
  add_cols(SELECTED = ifelse(GEN %in% selected$Code, "yes", "no")) %>% 
  filter(SELECTED == "yes") %>% 
  dplyr::summarise(mean_GY = mean(yield,na.rm = TRUE), n = n()) 

SD_blups<- as_tibble((blups_sel_mean2$mean_GY/mean(blups_sel2$BLUPs_sel, na.rm = T)) -1)*100
SD_WAASP<- as_tibble((waasb_sel_mean$mean_GY /mean(waasb_sel$WAASB_sel, na.rm = T)) -1)*100

SD_comb<- full_join(SD_blups, SD_WAASP, by = "value") %>% 
  dplyr::rename(SD_gain = value) %>% 
  tibble::add_column(Comp_name = c('BLUPs', 'WAASB')) %>% 
  relocate(Comp_name)

SD_comb$n_selected<- blups_sel_mean2$n
SD_comb
```


```{r mod metan stab4, warning=FALSE, message=TRUE, results='asis'}

blups_sel2$mean_blup <- mean(blups_sel2$BLUPs_sel, na.rm = T)
waasb_sel$mean_waasb <- mean(waasb_sel$WAASB_sel, na.rm = T)

#str(waasb_sel)
data_comb<- merge(blups_sel2, waasb_sel, by = c("GEN", "SELECTED"))
#names(data_comb)
## SD for each genotype
data_sel_perc <- data_comb %>%
 rowwise %>%
  mutate(Perc_blup_gain = ((BLUPs_sel/mean_blup)*100)-100) %>% 
  mutate(Perc_WAASB_gain = ((WAASB_sel/mean_waasb)*100)-100) %>% 
  as_tibble()

# data_sel_perc_mean <- data_sel_perc %>% 
#   dplyr::filter(SELECTED  == "yes")
# 
# mean(data_sel_perc_mean$Perc_blup_gain)

if (knitr::is_html_output()) {
  
print_table(data_sel_perc)
  
}else{
  
data_sel_perc[,1:7]
}


data_sel_perc<- data_sel_perc %>% 
  dplyr::relocate(GEN,SELECTED,BLUPs_sel,mean_blup,Perc_blup_gain,
                 WAASB_sel,mean_waasb ,Perc_WAASB_gain)

#write.xlsx(data_sel_perc, "./data/sel_SD_bb_2.xlsx")

data_sel_perc2 <- data_sel_perc %>% 
  dplyr::select(GEN,SELECTED, BLUPs_sel, WAASB_sel, Perc_blup_gain, Perc_WAASB_gain)

data_sel_perc2

```


```{r mod metan stab5, warning=FALSE, message=TRUE, results='asis'}
##BLUPs indexes
stab_blups<- blup_indexes(waasb_model_allMKT) 
stab_blups<- as_tibble(stab_blups$yield)

data_waasby <- waasb_model_allMKT$yield$model %>% 
  dplyr::filter(type != "ENV") %>% 
  dplyr::select("Code", "WAASBY", "OrWAASBY") %>% 
  dplyr::rename(GEN = Code)

stab_blups<- stab_blups %>% 
  full_join(data_waasby, by = "GEN")

if (knitr::is_html_output()) {
  
print_table(stab_blups)
  
}else{
  
stab_blups[,1:8]
}

# library(openxlsx)
# write.xlsx(stab_blups, "./data/blups_bb_2.xlsx")

```




#### Scenarios of waasby estimation


Planning different scenarios of waasby estimation by changing the weights assigned to the stability and the mean performance according to Olivoto et al. (2019)

```{r mod metan all_mkt stab6, warning=FALSE, message=FALSE, fig.height=figheight, fig.width=figwidth, fig.align="center", dpi = 600}
scenarios <- wsmp(waasb_model_allMKT,progbar = F)

scen1<- plot(scenarios, type = 1) + 
  theme(axis.text.y = element_text(size=8))

scen2<-plot(scenarios, type = 2) +
  theme(axis.text.y = element_text(size=8))

arrange_ggplot(scen1, scen2,
  guides = "collect",
  legend.position = "right",
  tag_levels = "a",
  tag_prefix = "(",
  tag_suffix = ")")
```


#### Coincidence index of genotype selection

Computes the coincidence index (Hamblin and Zimmermann, 1986) as follows:

$$
{CI = \frac{A-C}{M-C}\times 100}
$$

where *A* is the number of selected genotypes common to different methods; 
*C* is the number of expected genotypes selected by chance; 
and *M* is the number of genotypes selected according to the selection intensity.



```{r mod metan stab7, warning=FALSE, message=FALSE, fig.height=figheight, fig.width=figwidth, fig.align="center", dpi = 600}
coinc_1 <- stab_blups %>% dplyr::select(GEN,HMRPGV_R) %>% arrange(HMRPGV_R)
coinc_2 <- stab_blups %>% dplyr::select(GEN,RPGV_R) %>% arrange(RPGV_R)
coinc_3 <- stab_blups %>% dplyr::select(GEN,HMGV_R) %>% arrange(HMGV_R)
coinc_4 <- stab_blups %>% dplyr::select(GEN,OrWAASBY) %>% arrange(OrWAASBY)
coinc_5 <- stab_blups %>% dplyr::select(GEN,WAASB_R) %>% arrange(WAASB_R)

selc_perc<- round(nrow(stab_blups)*0.2)

coinc_1.1 <-1
coinc_1.2 <- coincidence_index(sel1 = coinc_1$GEN[1:selc_perc], 
                                        sel2 = coinc_2$GEN[1:selc_perc], 
                                        total = 72)/100
coinc_1.3 <- coincidence_index(sel1 = coinc_1$GEN[1:selc_perc], 
                                        sel2 = coinc_3$GEN[1:selc_perc], 
                                        total = 72)/100
coinc_1.4 <- coincidence_index(sel1 = coinc_1$GEN[1:selc_perc], 
                                        sel2 = coinc_4$GEN[1:selc_perc], 
                                        total = 72)/100
coinc_1.5 <- coincidence_index(sel1 = coinc_1$GEN[1:selc_perc], 
                                        sel2 = coinc_5$GEN[1:selc_perc], 
                                        total = 72)/100
coinc_2.2 <-1
coinc_2.3 <- coincidence_index(sel1 = coinc_2$GEN[1:selc_perc], 
                                        sel2 = coinc_3$GEN[1:selc_perc], 
                                        total = 72)/100
coinc_2.4 <- coincidence_index(sel1 = coinc_2$GEN[1:selc_perc], 
                                        sel2 = coinc_4$GEN[1:selc_perc], 
                                        total = 72)/100
coinc_2.5 <- coincidence_index(sel1 = coinc_2$GEN[1:selc_perc], 
                                        sel2 = coinc_5$GEN[1:selc_perc], 
                                        total = 72)/100
coinc_3.3<- 1
coinc_3.4 <- coincidence_index(sel1 = coinc_3$GEN[1:selc_perc], 
                                        sel2 = coinc_4$GEN[1:selc_perc], 
                                        total = 72)/100
coinc_3.5 <- coincidence_index(sel1 = coinc_3$GEN[1:selc_perc], 
                                        sel2 = coinc_5$GEN[1:selc_perc], 
                                        total = 72)/100
coinc_4.4 <- 1
coinc_4.5 <- coincidence_index(sel1 = coinc_4$GEN[1:selc_perc], 
                                        sel2 = coinc_5$GEN[1:selc_perc], 
                                        total = 72)/100
coinc_5.5 <- 1


coinc<- c(coinc_1.1,coinc_1.2,coinc_2.2,coinc_1.3,coinc_2.3,
          coinc_3.3,coinc_1.4, coinc_2.4, coinc_3.4,
          coinc_4.4, coinc_1.5, coinc_2.5,
          coinc_3.5, coinc_4.5,
          coinc_5.5)
  
z=matrix(0,5,5)
z[upper.tri(z)| row(z)==col(z)] <- coinc

rownames(z)=c(
"HMRPGV",
"RPGV",
'HMGV',
'WAASBY',
'WAASB')

colnames(z)=rownames(z)

plot<- ggcorrplot(z, colors = c("#6D9EC1", "gray" ,"#E46726"),  
           show.legend = T,
legend.title = "CI" ,lab_size=5,tl.srt = 90,type = c("upper"), lab = T,digits = 4,
outline.color = "white",pch.col = "white", tl.col = "blue",show.diag = F) +
  labs(title = "BLUP-based stability indexes coincidence across all market classes",
           subtitle = "Selection intensity of 20% top genotypes")

print(plot)

```


